Star-galaxy Classification Using Deep Convolutional Neural Networks

### 2016-08-14 paper/main.tex
update paper

\usepackage{graphicx}	% Including figure files
%\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage[ansinew]{inputenc}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{natbib}
%\usepackage{url}
\usepackage{hyperref}
\usepackage{placeins}
\usepackage{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}

@@

\changed{
In this section, we first provide a brief overview of deep neural networks
including convolutional neural networks.
We then describe the precise network architecture and the details of key
hyperparamters.
}

@@

To find the weights $\bmath{w}$ and biases $\bmath{b}$ which minimize the loss,
we use a technique called \textit{gradient descent},
where we use the following rules to update the parameters in each layer $l$:
\begin{align}
  \bmath{w}_l &\rightarrow
  \bmath{w}_l^{\prime}
  = \bmath{w}_l - \eta \frac{\partial L}{\partial \bmath{w}_l} \nonumber \\
  \bmath{b}_l &\rightarrow
  \bmath{b}_l^{\prime}
  = \bmath{b}_l - \eta \frac{\partial L}{\partial \bmath{b}_l},
  \label{eq:gradient_descent}
\end{align}
where $\eta$ is a small, positive number known as the \textit{learning rate}.
\changed{
The gradients can be computed using the backpropagation
procedure~\citep{rumelhart1988learning}.
A common approach to speed up training is to split the training data
into mini-batches~\citep{lecun1998efficient}.
In mini-batch gradient descent, instead of computing the gradients in
Equation~\ref{eq:gradient_descent} for the entire training data,
we only compute the gradient of randomly chosen training examples at each step.
As training examples are usually correlated,
the gradient computed from each mini-batch is a good approximation
of the overall gradient~\citep{bottou1998online}.
As a result, mini-batch gradient descent results in much faster convergence.
However, there is a trade-off: the lower the batch size is,
the lower the convergence rate will be;
the higher the batch size is, the longer it will take to compute the gradient
at each step~\citep{bousquet2008tradeoffs}.
Thus, a moderate batch size, combined with a decaying learning rate, is
generally used in practice.
}

@@

To implement our model, we use Python and the Lasagne
library~\citep{dieleman2015lasagne}, which is built on top of
Theano~\citep{theano2016theano}.
The Theano library simplifies the use of GPU for computation, and using the GPU
allows about an order of magnitude faster training than using just the CPU.
We note that training our network takes about forty hours on an NVIDIA Tesla
K40 GPU.
In the interest of scientific reproducibility, we make all our code available
at \href{https://github.com/EdwardJKim/dl4astro}{https://github.com/EdwardJKim/dl4astro}.

@@

To make final classifications,
we use our ConvNet model to make 64 sets of predictions for
64 transformations of the input images:
4 rotations, 4 horizontal translations, and 4 vertical translations
(with random horizontal reflections).
Although we use an identical network architecture for all transformations,
we consider each set of predictions as separate results from different models.
Finally, we use a model combination technique known as
Bayesian Model Combination~\citep[BMC;][]{Monteith2011}, which
uses Bayesian principles to generate an ensemble combination of different models.
Although the data augmentation step in Section~\ref{sec:data_augmentation} should
make our ConvNet model invariant to these types of transformations,
we find that applying BMC still results in a significant increase in performance.
\changed{
For a thorough discussion of BMC, we refer the reader to \citet{Monteith2011}
(See also \citet{carrascokind2014exhausting} for application of BMC to
photometric estimation, and \citet{kim2015hybrid} for combining star-galaxy
classifiers).
}

@@

\changed{
Following \citet{kim2015hybrid}, we use performance metrics that are well-suited
for probabilistic classifiers: Area Under the Curve (AUC)
for the Receiver Operating Characteristic (ROC) curve, completeness and purity,
and the Mean Squared Error (MSE).
A good probabilistic classifier should also provide well-calibrated
posterior probabilities.
Thus, to evaluate calibration performance,
we also use the calibration error and the absolute error in the estimation
of number of galaxies as performance metrics.
}

@@

\begin{table*}
  \caption{
    A summary of the classification performance metrics
    as applied to the CFHTLenS data.
    The definition of the metrics is summarized in Table~\ref{table:metrics}.
    The bold entries highlight the best performance values within each column.
\changed{
    Note that $p_{g}(c_g=0.96)$ and $c_{s}(p_s=0.97)$ require adjusting
    threshold values (\ie probability cuts), while other metrics do not.
    To obtain a galaxy completeness of $c_g=0.96$, we choose the threshold values
    0.9972, 0.9963, and 0.9927 for ConvNet, TPC${}_{\rmn{morph}}$, TPC${}_{\rmn{phot}}$,
    respectively;
    for star purity $p_s=0.97$, we choose 0.6990, 0.5297, and 0.8570 for
    ConvNet, TPC${}_{\rmn{morph}}$, TPC${}_{\rmn{phot}}$, respectively.
}
  }
  \centering
  \begin{tabular}{l c c c c c c c}
    \hline
    classifier & AUC & MSE & $p_{g}(c_g=0.96)$ & $c_{s}(p_s=0.97)$ & CAL & $ |\Delta N_g|/N_g$ &
\changed{
log loss
}
\\
    \hline
    ConvNet                       & \textbf{0.9948} & 0.0112          & \textbf{0.9972} & 0.8971          & \textbf{0.0197}  & \textbf{0.0029} &
\changed{
\textbf{0.0441}
}
\\
    TPC${}_{\rmn{morph}}$         & 0.9924          & \textbf{0.0109} & 0.9963          & \textbf{0.9268} & 0.0245           & 0.0056 &
\changed{
0.0809
}
\\
    TPC${}_{\rmn{phot}}$          & 0.9876          & 0.0189          & 0.9927          & 0.8044          & 0.0266           & 0.0101 &
\changed{
0.1085
}
\\
    \hline
  \end{tabular}
  \label{table:clens_metrics}
\end{table*}

@@

We present in Table~\ref{table:clens_metrics} a summary of the results obtained
by applying ConvNet, $\rmn{TPC}_{\rmn{morph}}$, and $\rmn{TPC}_{\rmn{phot}}$
on the test set of the CFHTLenS data.
The bold entries highlight the best technique for any particular metric.
ConvNet outperforms $\rmn{TPC}_{\rmn{morph}}$ in \changed{five} metrics
(AUC, $p_g$, CAL, $|\Delta N_g|/N_g$, \changed{and log loss}),
while $\rmn{TPC}_{\rmn{morph}}$ performs better in two metrics (MSE and $c_g$).
It is not surprising that $\rmn{TPC}_{\rmn{phot}}$, which is trained on only
magnitudes and colors, performs significantly worse than both ConvNet and
$\rmn{TPC}_{\rmn{phot}}$.
\sout{ As Figure~\ref{fig:clens_color_color} shows, there is significant overlap between stars and galaxies in color-color space.  }
Thus, magnitudes and colors alone are not sufficient to separate stars from
galaxies,
and morphology is critical in separating stars from galaxies.
ConvNet is able to learn the morphological features automatically from
the images, and the performance of ConvNet is therefore comparable to
that of $\rmn{TPC}_{\rmn{morph}}$,
which is trained on both morphological and photometric attributes.

@@

\begin{table*}
  \caption{
    A summary of the classification performance metrics
    as applied to the SDSS data.
\changed{
    To obtain a galaxy completeness of $c_g=0.96$, we choose the threshold values
    0.7558, 0.9989, and 0.9360 for ConvNet, TPC${}_{\rmn{morph}}$, TPC${}_{\rmn{phot}}$,
    respectively;
    for star purity $p_s=0.97$, we choose 0.6046, 0.0547, and 0.7449 for
    ConvNet, TPC${}_{\rmn{morph}}$, TPC${}_{\rmn{phot}}$, respectively.
}
  }
  \centering
  \begin{tabular}{l c c c c c c c}
    \hline
    classifier & AUC & MSE & $p_{g}(c_g=0.96) $ & $c_{s}(p_s=0.97)$ & CAL & $ |\Delta N_g|/N_g$ & \changed{log loss} \\
    \hline
    ConvNet               & 0.9952          & 0.0182          & 0.9915          & 0.9500          & \textbf{0.0243} & 0.0157 & \changed{\textbf{0.0731}} \\
    TPC${}_{\rmn{morph}}$ & \textbf{0.9967} & \textbf{0.0099} & \textbf{0.9977} & \textbf{0.9810} & 0.0254          & \textbf{0.0044} & \changed{0.0914} \\
    TPC${}_{\rmn{phot}}$  & 0.9886          & 0.0283          & 0.9819          & 0.8879          & 0.0316          & 0.0160 & \changed{0.1372} \\
    \hline
  \end{tabular}
  \label{table:sdss_metrics}
\end{table*}

@@

\changed{
Furthermore, our ConvNet is a supervised algorithm; and one of the criticisms
of supervised techniques is their difficulty in extrapolating past the limits
of available spectroscopic training data.
Since it is difficult to assess the classification performance without a deeper
spectroscopic sample, we have evaluated the performance on a test set
that is drawn from the same underlying distribution as the spectroscopic sample.
However, when our ConvNet model
---trained on sources from a spectroscopic sample---
is applied to a photometric sample
---which is often fainter than our training set--- 
the performance of ConvNet will be less reliable.
Combining our ConvNet model with unsupervised methods (\eg a template fitting
method) in this meta-classification framework will help improve the efficacy of
star-galaxy classification beyond the limits of spectroscopic training data.
}

@@

\changed{
The authors thank the referees for a careful reading of the manuscript
and comments that improved this work.
}
We acknowledge support from the 
National Science Foundation Grant No.\ AST-1313415.
RJB acknowledges support as an Associate
within the Center for Advanced Study at the University of Illinois.

@@

\changed{
This work has made use of the Theano library
(\href{http://deeplearning.net/software/theano/}{http://deeplearning.net/software/theano/}),
the Lasagne library
(\href{http://lasagne.readthedocs.io/)}{http://lasagne.readthedocs.io/}),
and the Nolearn library
(\href{http://pythonhosted.org/nolearn/}{http://pythonhosted.org/nolearn/}).
}

@@

\footnotesize{
\bibliographystyle{mnras}
\bibliography{sg_paper} }

>>>>

\usepackage{graphicx}	% Including figure files
%\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage[ansinew]{inputenc}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{natbib}
\usepackage{url}
%\usepackage{hyperref}
\usepackage{placeins}
\usepackage{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}

@@

\changed{
Neural networks have many hyperparameters, including those that specify the
network itself (\eg the size and non-linearity of each layer)
and those that specify how the network is trained
(\eg the mini-batch size or the learning rate).
Furthermore, the architecture of a neural network can have a
significant impact on its performance.
In this section, we provide a brief description of key hyperparameters in our
ConvNet model, and also present the network architecture.
}

@@

To find the weights $\bmath{w}$ and biases $\bmath{b}$ which minimize the loss,
we use a technique called \textit{gradient descent},
where we use the following rules to update the parameters in each layer $l$:
\begin{align}
  \bmath{w}_l &\rightarrow
  \bmath{w}_l^{\prime}
  = \bmath{w}_l - \eta \frac{\partial L}{\partial \bmath{w}_l} \nonumber \\
  \bmath{b}_l &\rightarrow
  \bmath{b}_l^{\prime}
  = \bmath{b}_l - \eta \frac{\partial L}{\partial \bmath{b}_l},
  \label{eq:gradient_descent}
\end{align}
where $\eta$ is a small, positive number known as the \textit{learning rate}.
\changed{
The gradients can be computed using the backpropagation
procedure~\citep{rumelhart1988learning}.
A common approach to speed up training is to split the training data
into mini-batches~\citep{lecun1998efficient}.
In mini-batch gradient descent, instead of computing the gradients in
Equation~\ref{eq:gradient_descent} for the entire training data,
we only compute the gradient of randomly chosen training examples at each step.
As training examples are usually correlated,
the gradient computed from each mini-batch is a good approximation
of the overall gradient~\citep{bottou1998online}.
As a result, mini-batch gradient descent results in much faster convergence.
However, there is a trade-off: the lower the batch size is,
the lower the convergence rate will be;
the higher the batch size is, the longer it will take to compute the gradient
at each step~\citep{bousquet2008tradeoffs}.
Thus, a moderate batch size, combined with a decaying learning rate, is
generally used in practice. We use a batch size of 128 in this work.
}

@@

To implement our model, we use Python and the Lasagne
library~\citep{dieleman2015lasagne}, which is built on top of
Theano~\citep{theano2016theano}.
The Theano library simplifies the use of GPU for computation, and using the GPU
allows about an order of magnitude faster training than using just the CPU.
We note that training our network takes about forty hours on an NVIDIA Tesla
K40 GPU.
In the interest of scientific reproducibility, we make all our code available
at \mbox{\url{https://github.com/EdwardJKim/dl4astro}}.

@@

To make final classifications,
we use our ConvNet model to make 64 sets of predictions for
64 transformations of the input images:
4 rotations, 4 horizontal translations, and 4 vertical translations
(with random horizontal reflections).
Although we use an identical network architecture for all transformations,
we consider each set of predictions as separate results from different models.
Finally, we use a model combination technique known as
Bayesian Model Combination~\citep[BMC;][]{Monteith2011}, which
uses Bayesian principles to generate an ensemble combination of different models.
Although the data augmentation step in Section~\ref{sec:data_augmentation} should
make our ConvNet model invariant to these types of transformations,
we find that applying BMC still results in a significant increase in performance.
\changed{
For a thorough discussion of BMC, we refer the reader to \citet{Monteith2011}
(See also \citet{carrascokind2014exhausting} for application of BMC to
photometric redshift estimation,
and \citet{kim2015hybrid} for combining star-galaxy classifiers).
}

@@

\changed{
Following \citet{kim2015hybrid}, we use performance metrics that are well-suited
for probabilistic classifiers: Area Under the Curve (AUC)
for the Receiver Operating Characteristic (ROC) curve, completeness and purity,
and the Mean Squared Error (MSE).
A good probabilistic classifier should also provide well-calibrated
posterior probabilities.
Thus, to evaluate calibration performance,
we also use the calibration error and the absolute error in the estimation
of number of galaxies.
}

@@

\begin{table*}
  \caption{
    A summary of the classification performance metrics
    as applied to the CFHTLenS data.
    The definition of the metrics is summarized in Table~\ref{table:metrics}.
    The bold entries highlight the best performance values within each column.
\changed{
    Note that $p_{g}(c_g=0.96)$ and $c_{s}(p_s=0.97)$ require adjusting
    threshold values (\ie probability cuts), while other metrics do not.
    To obtain a galaxy completeness of $c_g=0.96$, we choose the threshold values
    0.9972, 0.9963, and 0.9927 for ConvNet, TPC${}_{\rmn{morph}}$, and 
    TPC${}_{\rmn{phot}}$, respectively;
    for star purity $p_s=0.97$, we choose 0.6990, 0.5297, and 0.8570 for
    ConvNet, TPC${}_{\rmn{morph}}$, and TPC${}_{\rmn{phot}}$, respectively.
}
  }
  \centering
  \begin{tabular}{l c c c c c c c}
    \hline
    classifier & AUC & MSE & $p_{g}(c_g=0.96)$ & $c_{s}(p_s=0.97)$ & CAL & $ |\Delta N_g|/N_g$ &
\changed{
log loss
}
\\
    \hline
    ConvNet                       & \textbf{0.9948} & 0.0112          & \textbf{0.9972} & 0.8971          & \textbf{0.0197}  & \textbf{0.0029} &
\changed{
\textbf{0.0441}
}
\\
    TPC${}_{\rmn{morph}}$         & 0.9924          & \textbf{0.0109} & 0.9963          & \textbf{0.9268} & 0.0245           & 0.0056 &
\changed{
0.0809
}
\\
    TPC${}_{\rmn{phot}}$          & 0.9876          & 0.0189          & 0.9927          & 0.8044          & 0.0266           & 0.0101 &
\changed{
0.1085
}
\\
    \hline
  \end{tabular}
  \label{table:clens_metrics}
\end{table*}

@@

We present in Table~\ref{table:clens_metrics} a summary of the results obtained
by applying ConvNet, $\rmn{TPC}_{\rmn{morph}}$, and $\rmn{TPC}_{\rmn{phot}}$
on the test set of the CFHTLenS data.
The bold entries highlight the best technique for any particular metric.
ConvNet outperforms $\rmn{TPC}_{\rmn{morph}}$ in
\changed{
five
}
metrics
(AUC, $p_g$, CAL, $|\Delta N_g|/N_g$, 
\changed{
and log loss
}
),
while $\rmn{TPC}_{\rmn{morph}}$ performs better in two metrics (MSE and $c_g$).
It is not surprising that $\rmn{TPC}_{\rmn{phot}}$, which is trained on only
magnitudes and colors, performs significantly worse than both ConvNet and
$\rmn{TPC}_{\rmn{phot}}$.
\sout{ As Figure~\ref{fig:clens_color_color} shows, there is significant overlap between stars and galaxies in color-color space.  }
Thus, magnitudes and colors alone are not sufficient to separate stars from
galaxies,
and morphology is critical in separating stars from galaxies.
ConvNet is able to learn the morphological features automatically from
the images, and the performance of ConvNet is therefore comparable to
that of $\rmn{TPC}_{\rmn{morph}}$,
which is trained on both morphological and photometric attributes.

@@

\begin{table*}
  \caption{
    A summary of the classification performance metrics
    as applied to the SDSS data.
\changed{
    To obtain a galaxy completeness of $c_g=0.96$, we choose the threshold values
    0.7558, 0.9989, and 0.9360 for ConvNet, TPC${}_{\rmn{morph}}$, and
    TPC${}_{\rmn{phot}}$, respectively;
    for star purity $p_s=0.97$, we choose 0.6046, 0.0547, and 0.7449 for
    ConvNet, TPC${}_{\rmn{morph}}$, and TPC${}_{\rmn{phot}}$, respectively.
}
  }
  \centering
  \begin{tabular}{l c c c c c c c}
    \hline
    classifier & AUC & MSE & $p_{g}(c_g=0.96) $ & $c_{s}(p_s=0.97)$ & CAL & $ |\Delta N_g|/N_g$ &
\changed{
log loss
}
\\
    \hline
    ConvNet               & 0.9952          & 0.0182          & 0.9915          & 0.9500          & \textbf{0.0243} & 0.0157 &
\changed{
\textbf{0.0731}
}
\\
    TPC${}_{\rmn{morph}}$ & \textbf{0.9967} & \textbf{0.0099} & \textbf{0.9977} & \textbf{0.9810} & 0.0254          & \textbf{0.0044} & 
\changed{
0.0914
}
\\
    TPC${}_{\rmn{phot}}$  & 0.9886          & 0.0283          & 0.9819          & 0.8879          & 0.0316          & 0.0160 & 
\changed{
0.1372
}
\\
    \hline
  \end{tabular}
  \label{table:sdss_metrics}
\end{table*}

@@

\changed{
Our ConvNet model is a supervised algorithm, and one of the criticisms
of supervised techniques is their difficulty in extrapolating past the limits
of available spectroscopic training data.
Since it is difficult to assess the classification performance without a deeper
spectroscopic sample, we evaluated the performance using a test set
that is drawn from the same underlying distribution as the spectroscopic sample.
However, when our ConvNet model
---trained on sources from a spectroscopic sample---
is applied to a photometric sample
---which is often fainter than our training set--- 
the performance of ConvNet will be less reliable.
Combining our ConvNet model with unsupervised methods (\eg a template fitting
method) in the aforementioned meta-classification framework will
help improve the efficacy of star-galaxy classification
beyond the limits of spectroscopic training data.
}

@@

We acknowledge support from the 
National Science Foundation Grant No.\ AST-1313415.
RJB acknowledges support as an Associate
within the Center for Advanced Study at the University of Illinois.

@@

\changed{
This work has made use of the Theano library
(http://deeplearning.net/software/theano/),
the Lasagne library
(http://lasagne.readthedocs.io/),
and the Nolearn library
(http://pythonhosted.org/nolearn/).
}

@@

\bibliographystyle{mnras}
\bibliography{sg_paper}
